---
title: "Create Table with Sparklines"
subtitle: "Presentation at R Ladies Boston"
author: "Sharon Machlis"
date: "2021-01-12"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import data I previously downloaded from CovidActNow.org

```{r}
pacman::p_load(rio, dplyr, DT, sparkline)
latest_data <- import("data/latest_covid_by_state.csv") %>%
  select(state, newCasesPer100K = metrics.caseDensity, testPositivity = metrics.testPositivityRatio, lastUpdatedDate) %>%
  mutate(
    newCasesPer100K = round(newCasesPer100K, 1),
    testPositivity = round(testPositivity, 3)
  ) %>%
  arrange(desc(newCasesPer100K))

all_data <- import("data/all_covid_by_state.csv") %>%
  select(state, newCasesPer100K = metrics.caseDensity, testPositivity = metrics.testPositivityRatio, date) %>%
  filter(date >= as.Date("2020-04-01")) %>%
  mutate(
    newCasesPer100K = round(newCasesPer100K, 1),
    testPositivity = round(testPositivity, 3)
  ) %>%
  arrange(state, date)


glimpse(all_data)

```

```{r}
glimpse(latest_data)
```


## Prep data for sparkline column newCasesPer100K

I will brute-force reduce number of entries, because every day since March 1 will be a _lot_ of data points for a tiny sparkline. I'll start April 1 and take every 7th item only (once per week, same day each week). 

Then I'll use the sparkline package's `spk_chr()` function to generate the HTML for a sparkline line graph:

```{r, message=FALSE}
sparkline_data <- all_data %>%
  filter(date >= as.Date("2020-04-01")) %>%
  slice(which(row_number() %% 7 == 1)) %>%
  group_by(state) %>%
  summarize(
    CasesPer100KTrend = spk_chr(
      newCasesPer100K, type ="line",
      chartRangeMin = 0, chartRangeMax = max(newCasesPer100K, na.rm = TRUE)
    )
  )


```

I want to add my sparkline data frame's CasesPer100KTrend column to my latest_data data frame. I'll do that with a dplyr left_join to create a `table_data` data frame:

```{r}
table_data <- left_join(sparkline_data, latest_data)
glimpse(table_data)

```

This isn't really _necessary_, but I'm adding a region column to table_data so we'll have something interesting to filter by besides state names. I have a CSV file with state names and their US Census regions.

```{r}
regions <- rio::import("data/census_regions.csv") %>%
  select(state = `State Code`, region = Division)
glimpse(regions)
```

```{r}
table_data <- left_join(table_data, regions)
glimpse(table_data)

```




Create table

```{r}
datatable(table_data, escape = FALSE, 
          options = list(pageLength = 25,
            search = list(regex = TRUE),                                      fnDrawCallback = htmlwidgets::JS(
                                      '
function(){
  HTMLWidgets.staticRender();
}
'
)), rownames = FALSE, filter = 'top') %>%
      spk_add_deps()   %>%
      formatPercentage("testPositivity", digits = 1)
    

  



```








